### 2020-06-12 chapters/sota.tex
DeepAnT: A Deep Learning Approach for Unsupervised Anomaly Detection in Time Series

\subsection*{Univariate point outlier detection}

@@

\subsection{Neighbor-Based Pattern Detection for Windows Over Streaming Data: the Extra-N algorithm} \label{alg:extran}

@@

\subsection{Real-Time Stream Anomaly Detection with Hierarchical Temporal Memory}
Ahmad \& Purdy, 2016, present an anomaly detection technique based on an online algorithm called Hierarchical Temporal Memory (HTM)\footnote{https://github.com/numenta/nupic} \cite{Ahmad-HTM}. HTM has a few properties that are desirable for data stream anomaly detection scenarios. First, it is an unsupervised algorithm that requires no labels. This property comes in handy when monitoring a data stream in real-time, where labels for incoming data will likely be absent. Additionally, HTM adapts to noisy domains. In other words, random variations in the underlying distribution of incoming data that do not persist throughout time will be ignored. Furthermore, HTM detects both spatial and temporal anomalies --- \textit{i.e.} changes in magnitudes of data or unusual timing for particular patterns, respectively.

@@

\subsubsection*{Method classification}
Similarly to Extra-N presented in Section \ref{alg:extran}, the Hierarchical Temporal Memory (HTM) algorithm accepts both uni and multivariate time-series as input data but does not perform a multivariate analysis. Hence, we say it accepts both uni and multivariate inputs but is univariate in nature. HTM focuses on finding point outliers, but unlike Extra-N, which is density-based, HTM is a model-based approach. Specifically, HTM is a prediction model-based approach for univariate point outlier detection.

@@

\subsection{DeepAnT}
\subsubsection*{Method classification}
\subsubsection*{Applicability to our Hypothesis}

\iffalse
DeepAnT\\
point outlier, model-based, prediction\\
black box DL algorithm\\
explain properties\\
use what authors say in their review\\
\fi

\subsection*{Univariate subsequence outlier detection}
Similarly to point outlier detection methods, the authors split subsequence outlier detection methods into uni and multivariate ones (Figure \ref{fig:outlier-taxonomy}). Furthermore, the authors split univariate subsequence outlier detection methods into five different axes as seen in Figure \ref{fig:subsequence-outlier-taxonomy}.

@@

\subsection{A two window paradigm algorithm for change detection}
\subsubsection*{Method classification}
\subsubsection*{Time complexity}
\subsubsection*{Space complexity}
\subsubsection*{Applicability to our Hypothesis}

>>>>

\subsection{Univariate point outlier detection}

@@

\subsubsection{Neighbor-Based Pattern Detection for Windows Over Streaming Data: the Extra-N algorithm} \label{alg:extran}

@@

\subsubsection{Real-Time Stream Anomaly Detection with Hierarchical Temporal Memory}
Ahmad \& Purdy present an anomaly detection technique based on an online algorithm called Hierarchical Temporal Memory (HTM)\footnote{https://github.com/numenta/nupic} \cite{Ahmad-HTM}. HTM has a few properties that are desirable for data stream anomaly detection scenarios. First, it is an unsupervised algorithm that requires no labels. This property comes in handy when monitoring a data stream in real-time, where labels for incoming data will likely be absent. Additionally, HTM adapts to noisy domains. In other words, random variations in the underlying distribution of incoming data that do not persist throughout time will be ignored. Furthermore, HTM detects both spatial and temporal anomalies --- \textit{i.e.} changes in magnitudes of data or unusual timing for particular patterns, respectively.

@@

\subsubsection*{Method classification}
Similarly to Extra-N presented in Section \ref{alg:extran}, the Hierarchical Temporal Memory (HTM) algorithm accepts both uni and multivariate time-series as input data but does not perform a multivariate analysis. Hence, we say it accepts both uni and multivariate inputs but is univariate in nature. HTM focuses on finding point outliers, but unlike Extra-N, which is density-based, HTM is a model-based approach. Specifically, HTM is a prediction model-based approach for univariate point outlier detection (Figures \ref{fig:outlier-taxonomy} and \ref{fig:point-outlier-taxonomy}).

@@

\subsubsection{DeepAnT}
Munir et. al present DeepAnT, a deep learning-based approach for the detection of anomalies in time series data \cite{Munir-DeepAnT}, that uses unlabeled data to capture and learn the distribution of the data stream used to forecast the normal behavior of the time-series. It is unsupervised - \textit{i.e.} it requires no labels - which is an advantage in a streaming scenario where real-time decisions have to be made before labels are effectively-known. It is capable of detecting point anomalies in time series data with periodic and seasonal characteristics.

DeepAnT employs a Convolutional Neural Network (CNN) as its time-series predictor module. Before monitoring a data stream, the CNN model must be trained. The model can be trained with a small data set while achieving good generalization capabilities due to the effective parameter sharing of the CNN. Additionally, the data set used to train the model thus not require labels.

\subsubsection*{Method classification}
DeepAnT takes as input a uni or multivariate time-series but is a univariate method by nature --- \textit{i.e.} it does not find correlations between multiple time-dependent variables. The model used by DeepAnT is a Convolutional Neural Network (CNN), used to make predictions on the future time-series values and measure differences with the actual observed values. Hence, DeepAnT is a prediction model-based univariate point outlier detection method (Figures \ref{fig:outlier-taxonomy} and \ref{fig:point-outlier-taxonomy}).

\subsubsection*{Applicability to our Hypothesis}
Similar to the HTM algorithm presented, DeepAnT adapts to the underlying distribution of the data stream. In other words, if the data patterns shift from the initial configuration but stabilize, DeepAnT will accept it as it is. As mentioned, our goal is to build a monitoring system that alerts data pattern shifts continuously, until they match the initial configuration again. Furthermore, DeepAnT uses deep learning which is actually a subset of machine learning and just as inexplicable. In summary, we do not deem DeepAnT usable for our use case.

\subsection{Univariate subsequence outlier detection}
Similarly to point outlier detection methods, the authors split subsequence outlier detection methods into uni and multivariate ones (Figure \ref{fig:outlier-taxonomy}). Furthermore, the authors split univariate subsequence outlier detection methods into five different axes as seen in Figure \ref{fig:subsequence-outlier-taxonomy}.

@@

\subsubsection{A two window paradigm algorithm for change detection}
\subsubsection*{Method classification}
\subsubsection*{Time complexity}
\subsubsection*{Space complexity}
\subsubsection*{Applicability to our Hypothesis}
